Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy

### 2011-06-03 introduction.tex
Don't use shading; can't tell confidence intervals apart
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@883 ccdd05d5-56c1-461a-bf98-85f6876523b4

To this end, we have the ambition of reporting gravitational-wave candidates not minutes
\emph{after} the merger, but seconds \emph{before}.  By
looking for threshold crossings before the gravitational-wave signal leaves
the detection band, it is possible to trade some signal to noise ratio (\SNR{})
for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
for NS--NS binaries in Advanced \LIGO\ assuming the event rate predictions
in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
be detectable $t$ seconds before coalescence.  The solid line is the most
likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
shaded region corresponds to the confidence interval $\dot N_{\mathrm{low}}$ to $\dot 
N_{\mathrm{high}}$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
but at the cost of fewer total detectable events above an \SNR\ of 8.}
\end{center}
\end{figure}

>>>>

To this end, we have the ambition of reporting gravitational-wave candidates not minutes
\emph{after} the merger, but seconds \emph{before}.  By
looking for threshold crossings before the gravitational-wave signal leaves
the detection band, it is possible to trade some signal to noise ratio (\SNR{})
for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
for NS--NS binaries in Advanced \LIGO\ assuming the event rate predictions
in~\cite{Abadie:2010p10836}.
%
\begin{figure}
\begin{center}
\includegraphics{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
be detectable $t$ seconds before coalescence.  The heavy solid line is the most
likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
light solid lines represent the confidence interval $[\dot N_{\mathrm{low}}, \dot 
N_{\mathrm{high}}]$ described in
\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 
\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
but at the cost of fewer total detectable events above an \SNR\ of 8.}
\end{center}
\end{figure}
